---
title: "BIOS 500"
author: "Adrian Glover"
date: "2/15/2019"
output: html_document
---
Markdown
# Header
## subheader
plain text
*italics*
**bold**
'verbatim'
1. List
2. List
- Bullet 1
- Bullet 2
Footnote. [^1]
[^1]: The footnote.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Accessing and importing data

## Setting up a working directory
From: https://www.datacamp.com/community/tutorials/r-tutorial-read-excel-into-r

This tells you where your working directory is set currently:

```{r}
#getwd()
```

Then you can change it to the folder where your dataset is stored so that you don't have to type out full paths every time:

```{r}
#setwd("/Users/adrian/Desktop/")
```

## CSV files
From: https://www.lexjansen.com/phuse/2014/ad/AD04.pdf

To quickly copy a file or folder path to the clipboard in Mac OS X, select the file or folder in the OS X Finder, then hit Command+i to summon Get Info. Click and drag alongside “Where” to select the path, then hit Command+C to copy the full path to the clipboard.

You can also type '/' and then 'tab' and it will let you search:

```{r}
data1 <- read.csv("/Users/adrian/Desktop/work_countries_data.csv")
```

## Excel file
From: https://www.datacamp.com/community/tutorials/r-tutorial-read-excel-into-r and https://www.statmethods.net/input/importingdata.html 

If you save your data set as a tab-delimited texts file or '*.txt' file:

```{r}
data2 <- read.table("<FileName.txt>", header = TRUE, sep = ",", row.names = "id")
```

Or you can access the Excel files directly (the first row should contain the variable/column names) from https://rstudio-pubs-static.s3.amazonaws.com/413264_1a9480fe7a0448e0840d6cdb39b4ded2.html:

```{r}
library(readxl)
data3 <- read_excel("/Users/adrian/Desktop/GeorgiaLatinoDiabetesProgram.xlsx", col_names = TRUE)
data3
```

## SAS file
from https://www.datacamp.com/community/tutorials/r-data-import-tutorial#sas:

```{r}
install.packages("sas7bdat")
library(sas7bdat)
data4 <- read.sas7bdat("/Users/adrian/Desktop/DRAWERS/SASUniversityEdition/myfolders/SASCOURSE/data/education_income.sas7bdat")
data4
```

# Exploring data

## Cleaning data
See 'IDENTIFYING MISSING VALUES, DATA MODIFICATION' in https://www.lexjansen.com/phuse/2014/ad/AD04.pdf 

## Viewing data

```{r}
str(data3) # similar to proc contents

data3 # similar to proc print
```
 ## Exploring data
 https://dabblingwithdata.wordpress.com/2018/01/02/my-favourite-r-package-for-summarising-data/
 
 Frequencies:
 
```{r}

```
 
 
 Descriptive statistics: 
 
```{r}
# mean,median,25th and 75th quartiles,min,max
summary(data3, na.omit(data3))

# n, nmiss, unique, mean, 5,10,25,50,75,90,95th percentiles 
# 5 lowest and 5 highest scores
install.packages("Hmisc")
library(Hmisc)
describe(data3)

# item name ,item number, nvalid, mean, sd, 
# median, mad, min, max, skew, kurtosis, se
install.packages("psych")
library(psych)
describe(data3)

# creating tables with Kable
# mean, stdev, min, med, max, MAD, IQR, CV, skewness, SE.skewness, kurtosis, n.valid, pct.valid
install.packages("summarytools")
install.packages("knitr")
install.packages("kableExtra")
library(summarytools)
mydata <- summarytools::descr(data3)
mydata
View(mydata)
kableExtra::kable(as.data.frame(summarytools::descr(data3)))
# can also transpose to arrange variables by row instead
kableExtra::kable(as.data.frame(summarytools::descr(data3, transpose = TRUE)))
# there is also a fancier option that summarizes a whole dataframe that you can output directly into markdown or HTML
mydata2 <- dfSummary(data3)
View(mydata2)
```
 
Descriptive statistics by group:

```{r}
psych::describeBy(data3, "sexo")
```

Univariate statistics:
https://flowingdata.com/2012/05/15/how-to-visualize-and-compare-distributions/

http://sites.utexas.edu/sos/guided/descriptive/numericaldd/descriptiven2/groupedhist/

```{r}
library(psych)

# Univariate stats
describe(data3$yrs_school)

# Histogram
hist(data3$yrs_school, breaks = 5, right = T)

# Grouped histogram
yrs_school_male <- data3$yrs_school[data3$sexo==1]
yrs_school_female <- data3$yrs_school[data3$sexo==2]
hist(yrs_school_male, main = "Distribution of Yrs of School (Males)", xlab = "Yrs school", breaks = 5, right = T, xlim = c(1,6), ylim = c(0, 250), col = "Orange")
hist(yrs_school_female, main = "Distribution of Yrs of School (Females)", xlab = "Yrs school", breaks = 5, right = T, xlim = c(1,6), ylim = c(0, 250), col = "Orange")

# Boxplot
boxplot(data3$yrs_school)
```



