---
title: "Model Selection for MLR"
author: "Adrian Glover"
date: "3/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

## Homework Description

Lab 4 Homework

This assignment, worth 20 pts total.

The SAS dataset jet2(located in ‘S:\course\BIOS501\lab exercise data\’) contains data from a study on the appearance and technological factors of 22 new aircraft designs. For all technological factors, higher values correspond to more advanced technology. The following variables are included:

  FFD – Time of first appearance of aircraft design, measured in months after January 1940.
  Natural log transformed to satisfy normality assumption.

(Smaller values mean earlier appearances/older designs.)

  SPR – Specific power, proportional to power per unit weight
  RGF – Flight range factor
  PLF – Payload as a function of gross weight of the aircraft
  SLF – Sustained load factor
  CAR – Y = Aircraft can land on a carrier N = Aircraft cannot land on a carrier
 
Researchers are interested in:

  i) Predicting the age of a jet based on its technological specifications
  ii) Identifying any potential technological trends over time
 
Prior to conducting the regression analysis, you should do an exploratory analysis of each variable. You are not required to report any results on the HW, but include this in your SAS code with a brief comment on each variable. **

Recall that dependent response variables in PROC REG must be NUMERIC, i.e.,you must create a dummy variable if the response variable is CHARACTER.

# Author: Joan Braithwaite 
Activate the `sas7bdat` library  (after installing the package sas7bdat)

```{r}
library(sas7bdat)
```

Read in the SAS data  (using forward slashes in path)

```{r}
Jet <- read.sas7bdat("/Users/adrian/bios501/jet2.sas7bdat")
```

str(data.frame) = structure of data, provides data class, num of observations vs variables, and name of class of each variable and preview of its contents -- similar to PROC CONTENTS

view(data.frame) = opens and view the content of the data frame  -- similar to PROC PRINT

```{r}
str(Jet)
Jet
View(Jet)
head(Jet)
tail(Jet)
```

Exploratory data analysis
```{r}
summary(Jet)
pairs(Jet)
plot(Jet$FFD, Jet$SPR, main='FFD/SPR', xlab='FFD', ylab='SPR', col="red")
```

Create dummy variables for 'CAR' (binary Y/N variable)
from https://cran.r-project.org/web/packages/fastDummies/vignettes/making-dummy-variables.html

```{r}
install.packages("fastDummies")
library(fastDummies)
dummy_cols(Jet, remove_first_dummy = TRUE)
Jet2 <- data.frame(dummy_cols(Jet, remove_first_dummy = TRUE))
Jet2
```



All-possible-regressions strategy
```{r}
# Full model
fit <- lm(FFD ~ SPR + RGF + PLF + SLF + CAR_Y, data=Jet2)
summary(fit) # show results

# Other useful functions
coefficients(fit) # model coefficients
confint(fit, level=0.95) # CIs for model parameters 
fitted(fit) # predicted values
residuals(fit) # residuals
anova(fit) # anova table 
vcov(fit) # covariance matrix for model parameters 
influence(fit) # regression diagnostics
```
  
```{r}
library(MASS)
  y = Jet2$FFD
  x1 = Jet2$SPR
  x2 = Jet2$RGF
  x3 = Jet2$PLF
  x4 = Jet2$SLF
  x5 = Jet2$CAR_Y
fit <- lm(y ~ x1 + x2 + x3 + x4 + x5, data=Jet2)
step <- stepAIC(fit, direction="forward")
step <- stepAIC(fit, direction="backward")
step <- stepAIC(fit, direction="both")
step$anova # display results

# Final Model
fit <- lm(y ~ x1 + x2 + x4 + x5, data=Jet2)
fit
```

Forward selection strategy
```{r}
null <-lm(FFD~1, data=Jet2)  #No variables included in the model
full <- lm(FFD ~., data=Jet2) #Full Model with all variables

# This tells R to start with the null model and search through models lying in the
# range between the null and full model using the forward selection algorithm.

step(null, scope=list(lower=null, upper=full), direction="forward")
```

Backward elimination strategy
```{r}
step(full, data =Jet2, direction="backward")
```


Stepwise selection strategy
```{r}
step(null, scope = list(upper=full), data=Jet2, direction="both")
```

Run the full model
```{r}
fit <- lm(FFD ~ SPR + RGF + CAR + SLF, data=Jet2)
summary(fit) # show results
```
